\LARGE \bf Enriching Robot's Actions with Affective Movements

### 2017-06-11 RO-MAN-2017-2/RO-MAN2017-2.tex
Revised

\section{Introduction}
\input{./Parts/Introduction}
\section{Related Work}
\label{sec:related_work}
\input{./Parts/RelatedWork}
\section{Basic Concepts}
\label{sec:concepts}
The system is based on six main concepts that structure and enable the enrichment of actions with emotions. (i) \textit{Action message} establishes the structure of the message to describe any kind of action (i.e., simple and compound). This message also specifies how the actions are executed (i.e., in parallel or in sequence) and which action is predominant (i.e., primary or secondary). (ii) \textit{Emotional parameters} describe how the emotional enrichment should be done to convey a specific emotion in a specific simple action. This description could also include addition of other simple actions and vary over time. (iii) \textit{Character description} enables the possibility to establish how to modify emotional expressions to generate diverse "rhythms". (iv) \textit{Emotional Execution Tree} (EXT) is a computational representation of desired actions that should be executed. This tree is created from the action message description and then modified using the emotional parameters and character description. The remaining two concepts are explained below.

\subsection{Emotion Parameters}
The emotion parameters specify how actions parameters should be modified in order to express an emotion. These parameters could include a sequence of descriptors. For example, these descriptors could be used to generate changes of velocity over time or position to express an emotion. However, there is going to be a diverse set of parameters and descriptors depending on the nature of the action. For example, consider the actions \textit{move body} and \textit{speak}. The human action of speaking is related to changes in the vocal cords to produce different sounds, while \textit{move body} is directly related to the body movement. The speaking action would have the following parameters: text, pitch and tone. On the other hand, the action \textit{move body} would have as parameters the desired destination, velocity, and trajectory constraints. However, in order to express emotion in speak action, it is required to modify pitch variables (i.e. fundamental frequency level, contour and jitter), intensity and speech rate~\cite{Gunes2011}. On the other hand, changing linear velocity and adding oscillation could be used to convey emotion in the move body action~\cite{Angel2016c}. Since each action requires modification on diverse parameters, a variety of emotional parameters would be necessary to describe the modification of all type of actions. So, it was decided to focus on the parameters that could be used to express emotions with the robots we have considered: Triskarino and Keepon. It was determined that these parameters could just use movement actions, which are described by position, velocity and acceleration.    
\subsection{Simple and Compound Actions}
\input{./Parts/SimpleActions} 
\section{Emotional Tree}
\label{sec:emotional_execution_tree}
\input{./Parts/EmotionalTree}
\section{Describing Components}
\label{sec:description}
\input{./Parts/description}
\section{Implementation}
\label{sec:implementation}
\input{./Parts/Implementation}
\section{Conclusions and Further work}
An Emotional Enrichment System has been designed and implemented to enrich robots' movements with emotions. To achieve this,  an Emotional Execution Tree created from three different types of nodes was used: simple actions, parallel, and sequential. Simple actions are functions that map a set of parameters to specific movements. Sequential nodes execute the sequence of actions associated to this type of node in a sequential order, while parallel executes them all at the same time. To enable synchronization among simple actions, parallel nodes could be one of four different subtypes, and sequential just one of two different subtypes. A formalization of the Emotional Execution Tree and the principal considerations during the implementation of the system have also been provided. To show the system's versatility, it was used with quite different platforms such as a Keepon Pro and Triskarino. Keepon was used to perform simple actions, while Triskarino was used to test complex actions with different parametrizations of emotions. The system has been used in different case studies, in which has facilitate emotions descriptions.

>>>>

\section{Introduction}
\input{./Parts/Introduction}
\section{Related Work}
\label{sec:related_work}
\input{./Parts/RelatedWork}
\section{Basic Concepts}
\label{sec:concepts}
The system is based on six main concepts that structure and enable the enrichment of actions with emotions. (i) \textit{Action message} establishes the structure of the message to describe any kind of action (i.e., simple and compound). This message also specifies how the actions are executed (i.e., in parallel or in sequence) and which action is predominant (i.e., primary or secondary). (ii) \textit{Emotional parameters} describe how the emotional enrichment should be done to convey a specific emotion in a specific simple action. This description could also include addition of other simple actions and vary over time. (iii) \textit{Character description} enables the possibility to establish how to modify emotional expressions to generate diverse "rhythms". (iv) \textit{Emotional Execution Tree} (EXT) is a computational representation of desired actions that should be executed. This tree is created from the action message description and then modified using the emotional parameters and character description. The remaining two concepts rquire amore detailed description, reported below.

\subsection{Emotion Parameters}
The emotion parameters specify how actions parameters should be modified in order to express an emotion. These parameters could include a sequence of descriptors. For example, these descriptors could be used to generate changes of velocity or position over time to express an emotion. However, there is going to be a diverse set of parameters and descriptors depending on the nature of the action. For example, consider the actions \textit{move body} and \textit{speak}. The human action of speaking is related to changes in the vocal cords to produce different sounds, while \textit{move body} is directly related to the body movement. The speaking action would have the following parameters: text, pitch, and tone. On the other hand, the action \textit{move body} would have as parameters the desired destination, velocity, and trajectory constraints. However, in order to express emotion in speak action, it is required to modify pitch variables (i.e., fundamental frequency level, contour and jitter), intensity and speech rate~\cite{Gunes2011}. On the other hand, changing linear velocity and adding oscillation could be used to convey emotion in the move body action~\cite{Angel2016c}. Since each action requires modification of diverse parameters, a variety of emotional parameters would be necessary to describe the modification of all types of actions. So, it was decided to focus on the parameters that could be used to express emotions with the robots we have considered: Triskarino and Keepon. It was determined that these parameters could just use movement actions, which are described by position, velocity and acceleration.    
\subsection{Simple and Compound Actions}
\input{./Parts/SimpleActions} 
\section{Emotional Tree}
\label{sec:emotional_execution_tree}
\input{./Parts/EmotionalTree}
\section{Describing Components}
\label{sec:description}
\input{./Parts/description}
\section{Implementation}
\label{sec:implementation}
\input{./Parts/Implementation}
\section{Conclusions and Further work}
An Emotional Enrichment System has been designed and implemented to enrich robots' movements with emotions. To achieve this,  an Emotional Execution Tree created from three different types of nodes was used: simple actions, parallel, and sequential. Simple actions are functions that map a set of parameters to specific movements. Sequential nodes execute the sequence of actions associated to this type of node in a sequential order, while parallel executes them all at the same time. To enable synchronization among simple actions, parallel nodes could be one of four different subtypes, and sequential just one of two different subtypes. A formalization of the Emotional Execution Tree and the principal considerations during the implementation of the system have also been provided. To show the system's versatility, it was used with quite different platforms such as a Keepon Pro and Triskarino. Keepon was used to perform simple actions, while Triskarino was used to test complex actions with different parametrizations of emotions. The system has been used in different case studies, in which has facilitate emotions descriptions.
